Introducing an event-based architecture for concurrent and distributed evolutionary algorithms

### 2018-05-29 paper.tex
Spell check

\section{Introduction}
Cloud computing is increasingly becoming the dominant way of running
the server side of most enterprise applications nowadays, the same as the browser is
the standard platform for the client side. Besides the convenience of the pay-as-you-go
model, it also offers a way of describing the infrastructure as part
of the code, so that it is much easier to reproduce results and has 
been a boon for scientific computing.
%The availability of resources only limited by your budget, and in many
%cases for free, as well as the inherent reproducibility of cloud
%deployments, has been a boon for scientific computing. 
However,
programming the cloud means that monolithic applications, that is,
applications built on a single stack of services that communicate by
layers, are no longer an efficient architectural design for scientific workflows. Cloud
architectures favor asynchronous communication over heterogeneous %favor: american
resources, and shifting from mostly sequential and monolithic 
to an asynchronously parallel architecture will also imply important
reformulation of the algorithms in order to take full advantage of these technologies.
  Cloud-native applications add a layer of abstraction to the
  underlying distributed computing system, seamlessly integrating
  different elements in a single data flow, allowing the user to just focus on code and service connections. Services are {\em native} points
  in this new architecture, departing from a monolithic or even
  distributed paradigm to become a loosely collection of
services, in fact {\em microservices} \cite{microservices}, which in
 many cases are stateless, reacting to some event and
  {\em living} only while they are doing some kind of processing.
%   and not keeping state from one invocation to the next.
% redundant - Mario  
  Reactive systems not only allow massive scaling and independent deployment they are also more economical than other monolithic options.  Platform as a service (PaaS) or even Container as a Service (CaaS) approaches need to be running all
  the time in order to maintain their state, so they are paid for their size and
  time they remain active.
  At any rate, while one of the main selling points of Functions as a Service (FaaS) is their ultra-fast activation time, from our point of view their most interesting feature is the fact that they provide stateless processing. An important caveat
  of stateless processing is that algorithms must be adapted to this fact and
  turned, at least in part, into a series of stateless steps thaworking on a data stream.
    It is also taken to an atomic extreme
    with the so-called serverless architectures \cite{Varghese2018849},
    which allow vendors and users to deploy code as single, stateless
    functions, that get activated via {\em rules}, {\em triggers} or
    explicitly, reacting to events consumed from
    a message queue.
    %or bus that
    % generally reacting to events from a message queue or bus that
    % what is generated? rules are triggered and triggers fire in
    % response to events
    % Please change it to whatever you think it's the best - JJ
    %is the backbone of the application. 
    The first commercial
    implementation of this kind of architecture was released by Amazon
    with its Lambda product, to be closely followed by releases by Azure
    (Microsoft) and Google and OpenWhisk, an open source implementation
    released by IBM \cite{Baldini2016287}.

@@

  In this section we compare the performance of KafkEO against an implementation of
  the EvoSpace \cite{GValdez2015} pool-based architecture, using the first five
  functions of the Noiseless BBOB testbed \cite{hansen2016coco}, 
  which are real-parameter, single-objective, separable functions, namely: Sphere,,  ellipsoidal, which is highly multimodal,  Rastrigin,  Buche-Rastrigin,  and the purely lineal function called linear  slope. It is expected that the two algorithms achieve similar results as
  they are functionally equivalent. The EvoSpace implementation follows the
  basic EvoSpace model in which EvoWorkers asynchronously interact with the
  population pool by taking samples of the population to perform a standard
  evolutionary search on the samples, to then return newly evolved solutions back
  to the pool.

@@

  The evolutionary algorithm implemented in KafkEO used the same code,
  also delegating the evolutionary operations to the standard DEAP
  library, written in
  Python \cite{fortin2012deap},
%  The parameters for the evolutionary algorithm were fixed and
%  are shown in Table \ref{tab:GAparams}.
  using 12 for tournament size, a gaussian mutation sith sigma = 0.05 and a probability between 0.1 and 0.6, plus two point crossover with probability between 0.8 and 1; these are the default parameters.  
  % These parameters have been
  % chosen as the usual, and default, setup in any experiment, without
  % making any attempt to adapt it to this particular flavor of the
  % algorithm.
  In particular, the tournament size injects a high selective
  pressure which is known to decrease diversity. The system also allows
  to set different parameters for every instance; in this proof of concept
  only two parameters were randomly set, {\em Mutation Probability}
  uniformly random in the $[.1,.6]$ range, and {\em Crossover
    Probability} random on $[.8,1]$. This is one deviation from the
  standard evolutionary algorithm, but has been proved in the past to
  provide good results without needing to fine tune different
  parameters \cite{tanabe2013evaluation}.

>>>>

\section{Introduction}
Cloud computing is increasingly becoming the dominant way of running
the server side of most enterprise applications nowadays, the same as the browser is
the standard platform for the client side. Besides the convenience of the pay-as-you-go
model, it also offers a way of describing the infrastructure as part
of the code, so that it is much easier to reproduce results and has 
been a boon for scientific computing.
%The availability of resources only limited by your budget, and in many
%cases for free, as well as the inherent reproducibility of cloud
%deployments, has been a boon for scientific computing. 
However,
programming the cloud means that monolithic applications, that is,
applications built on a single stack of services that communicate by
layers, are no longer an efficient architectural design for scientific workflows. Cloud
architectures favor asynchronous communication over heterogeneous %favor: american
resources, and shifting from mostly sequential and monolithic 
to an asynchronously parallel architecture will also imply important
reformulation of the algorithms in order to take full advantage of these technologies.
  Cloud-native applications add a layer of abstraction to the
  underlying distributed computing system, seamlessly integrating
  different elements in a single data flow, allowing the user to just focus on code and service connections. Services are {\em native} points
  in this new architecture, departing from a monolithic or even
  distributed paradigm to become a loosely collection of
services, in fact {\em microservices} \cite{microservices}, which in
 many cases are stateless, reacting to some event and
  {\em living} only while they are doing some kind of processing.
%   and not keeping state from one invocation to the next.
% redundant - Mario  
  Reactive systems not only allow massive scaling and independent deployment they are also more economical than other monolithic options.  Platform as a service (PaaS) or even Container as a Service (CaaS) approaches need to be running all
  the time in order to maintain their state, so they are paid for their size and
  time they remain active.
  At any rate, while one of the main selling points of Functions as a Service (FaaS) is their ultra-fast activation time, from our point of view their most interesting feature is the fact that they provide stateless processing. An important caveat
  of stateless processing is that algorithms must be adapted to this fact and
  turned, at least in part, into a series of stateless steps working on a data stream.
    It is also taken to an atomic extreme
    with the so-called serverless architectures \cite{Varghese2018849},
    which allow vendors and users to deploy code as single, stateless
    functions, that get activated via {\em rules}, {\em triggers} or
    explicitly, reacting to events consumed from
    a message queue.
    %or bus that
    % generally reacting to events from a message queue or bus that
    % what is generated? rules are triggered and triggers fire in
    % response to events
    % Please change it to whatever you think it's the best - JJ
    %is the backbone of the application. 
    The first commercial
    implementation of this kind of architecture was released by Amazon
    with its Lambda product, to be closely followed by releases by Azure
    (Microsoft) and Google and OpenWhisk, an open source implementation
    released by IBM \cite{Baldini2016287}.

@@

  In this section we compare the performance of KafkEO against an implementation of
  the EvoSpace \cite{GValdez2015} pool-based architecture, using the first five
  functions of the Noiseless BBOB testbed \cite{hansen2016coco}, 
  which are real-parameter, single-objective, separable functions,
  namely: Sphere,  ellipsoidal, which is highly multimodal,
  Rastrigin,  Buche-Rastrigin,  and the purely lineal function called
  linear  slope. It is expected that the two algorithms achieve
  similar results as 
  they are functionally equivalent. The EvoSpace implementation follows the
  basic EvoSpace model in which EvoWorkers asynchronously interact with the
  population pool by taking samples of the population to perform a standard
  evolutionary search on the samples, to then return newly evolved solutions back
  to the pool.

@@

  The evolutionary algorithm implemented in KafkEO used the same code,
  also delegating the evolutionary operations to the standard DEAP
  library, written in
  Python \cite{fortin2012deap},
%  The parameters for the evolutionary algorithm were fixed and
%  are shown in Table \ref{tab:GAparams}.
  using 12 for tournament size, a Gaussian mutation with sigma = 0.05 and a probability between 0.1 and 0.6, plus two point crossover with probability between 0.8 and 1; these are the default parameters.  
  % These parameters have been
  % chosen as the usual, and default, setup in any experiment, without
  % making any attempt to adapt it to this particular flavor of the
  % algorithm.
  In particular, the tournament size injects a high selective
  pressure which is known to decrease diversity. The system also allows
  to set different parameters for every instance; in this proof of concept
  only two parameters were randomly set, {\em Mutation Probability}
  uniformly random in the $[.1,.6]$ range, and {\em Crossover
    Probability} random on $[.8,1]$. This is one deviation from the
  standard evolutionary algorithm, but has been proved in the past to
  provide good results without needing to fine tune different
  parameters \cite{tanabe2013evaluation}.
